---
title: "Hyperparameters and Tuning"
author: Dr. Gilbert
format: 
  revealjs:
    smaller: true
date: today
date-format: long
theme: serif
incremental: true
fontsize: 22pt
---

```{r global-options, include=FALSE}
#Note -- for taller slides, set height option on revealjs to a large pixel number, ie. 900

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(patchwork)
library(kableExtra)
library(memer)
library(marginaleffects)
library(rpart.plot)

tidymodels_prefer()

ames_known_prices <- ames %>%
  filter(!is.na(Sale_Price))

set.seed(123)

ames_split <- initial_split(ames, prop = 0.9)

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

ames_folds <- vfold_cv(ames_train, v = 5)

options(kable_styling_bootstrap_options = c("hover", "striped"))
options(scipen = 999)

#Set ggplot base theme
theme_set(theme_bw(base_size = 18))

dt1_spec <- decision_tree(tree_depth = 1) %>%
  set_engine("rpart") %>%
  set_mode("regression")

dt_rec <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_other(all_nominal_predictors()) %>%
  step_impute_median(all_numeric_predictors())

dt1_wf <- workflow() %>%
  add_model(dt1_spec) %>%
  add_recipe(dt_rec)

dt1_fit <- dt1_wf %>%
  fit(ames_train)

dt8_spec <- decision_tree(tree_depth = 8, min_n = 2, cost_complexity = 0) %>%
  set_engine("rpart") %>%
  set_mode("regression")

dt8_wf <- workflow() %>%
  add_model(dt8_spec) %>%
  add_recipe(dt_rec)

dt8_fit <- dt8_wf %>%
  fit(ames_train)
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

## Motivation

. . .

We've discussed several new model classes recently

. . . 

With each class, we've been required to set model parameters manually, prior to the model seeing any data

. . .

**Ridge Regression and LASSO:**

::::{.columns}

:::{.column width="50%"}

+ `mixture`

:::

:::{.columns width="50%"}

+ `penalty`

:::

::::

. . .

**Nearest Neighbor Regressors:**

+ `neighbors`

. . . 

**Tree-Based Models:** 

+ `max_depth`

. . .

**Ensembles (Random Forest, Boosted Trees):**

::::{.columns}

:::{.column width="33%"}

+ `mtry`

:::

:::{.column width="33%"}

+ `trees`

:::

:::{.column width="33%"}

+ `learn_rate`

:::

::::

+ etc.

. . .

Every modeling choice we make should be justified...

. . .

How can we make justifiable choices for these parameters?

## Model Hyperparameters

. . . 

A *hyperparameter* is a model parameter (setting) that must be chosen prior to the model being fit to training data

. . . 

That is, *hyperparameters* are set outside of the model fitting process

. . . 

If these parameters need to be determined prior to seeing data, how can we justify (or even 🤞 optimize) our choices?

## Keep Calm and ...

## Keep Calm and Cross-Validate

. . . 

I told you that cross-validation was useful for more than just estimating more stable performance estimates

. . . 

With cross-validation, we can try multiple combinations of hyperparameter settings

. . . 

The cross-validation procedure results in performance estimates (and standard error estimates) for each model and hyperparameter combination

. . . 

From here, we can identify not only the best model, but also the best hyperparameter choices

. . . 

We'll call this process *tuning*

## Playing Along

. . . 

Again, we'll continue with the `ames` data

1. Open your notebook from our last two meetings
2. Run all of the existing code
3. Add a new section to your notebook on *hyperparameters and model tuning*

## Tuning a Decision Tree Regressor

. . . 

Decision tree regressors are a great first model class to tune because optimizing the `max_depth` parameter is intuitive

. . . 

The *deeper* a decision tree, the more *flexible* it is, and the more likely it is to overfit

```{r}
dt8_fit_for_plot <- dt8_fit %>%
  extract_fit_engine()

rpart.plot(dt8_fit_for_plot, roundint = FALSE)
```

## Tuning a Decision Tree Regressor

Decision tree regressors are a great first model class to tune because optimizing the `max_depth` parameter is intuitive

The *deeper* a decision tree, the more *flexible* it is, and the more likely it is to overfit

If a tree is *too shallow* though, it will be underfit

```{r}
dt1_fit_for_plot <- dt1_fit %>%
  extract_fit_engine()

rpart.plot(dt1_fit_for_plot, roundint = FALSE)
```
. . .

Let's tune a tree to identify the optimal depth for predicting selling prices of homes in the Ames, Iowa data

## Defining a Model with Tuning Parameters

. . . 

Most of model tuning is done exactly as we approached assessing models robustly with cross-validation

. . . 

The only difference is that we'll indicate parameters to be *tuned* by setting them to `tune()`

```{r}
#| echo: true
#| eval: true

dt_spec <- decision_tree(tree_depth = tune()) %>%
  set_engine("rpart") %>%
  set_mode("regression")

dt_rec <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_other(all_nominal_predictors()) %>%
  step_impute_median(all_numeric_predictors())

dt_wf <- workflow() %>%
  add_model(dt_spec) %>%
  add_recipe(dt_rec)
  
```

## Tuning a Model

. . . 

Now that we've defined our model, we are ready to tune it

. . . 

We can't use `fit()` or `fit_resamples()` for this -- we'll need to `tune_grid()` instead

. . .

The *grid* in `tune_grid()` refers to the fact that we'll need to define a grid (collection) of hyperparameter combinations for the model to tune over

. . . 

We can use the `grid_regular()` function to create a *regular* grid of sensible values to try for each hyperparameter

. . . 

```{r}
#| echo: true
#| eval: true

tree_grid <- grid_regular(
  tree_depth(),
  levels = 10
)
```

## A Note on Complexity

. . . 

It is important to realize what we are asking for here... 

. . . 

The `levels` argument of `grid_regular()` determines the number of hyperparameter settings for *each* hyperparameter to be tuned

+ Since we are tuning just a single hyperparameter, we'll obtain just 10 settings
+ If we were tuning two hyperparameters, then we'd obtain $10\times 10 = 100$ combinations of hyperparameter settings!

. . . 

In addition to all of these hyperparameter settings, remember that we're using cross-validation to assess the resulting models

+ In this case, we have chosen to use 5-fold cross-validation

. . . 

All this together means that we are about to train and assess $10\times 5 = 50$ decision trees

. . . 

The more hyperparameters we tune / the finer our grid / the more cross-validation folds we utilize, the longer our tuning procedure will take

## Let's Tune!

. . . 

Okay, let's use cross validation to tune our decision tree and find an optimal choice for the maximum tree depth

. . . 

```{r}
#| eval: false
#| echo: true

dt_tune_results <- dt_wf %>%
  tune_grid(
    grid = tree_grid,
    resamples = ames_folds
  )

dt_tune_results %>%
  show_best(n = 10)
```

. . . 

```{r}
#| eval: true
#| echo: false

dt_tune_results <- dt_wf %>%
  tune_grid(
    grid = tree_grid,
    resamples = ames_folds
  )

dt_tune_results %>%
  show_best(n = 10) %>%
  kable() %>%
  kable_styling()
```

## Visualizing the Tuning Results

. . . 

We see that several tree depths result in models that perform similarly to one another

. . .

Eventually, including additional flexibility doesn't translate into improvements in our model performance metrics

. . . 

```{r}
#| echo: true
#| eval: true

dt_tune_results %>%
  autoplot()
```

## Finalizing the Workflow and Fitting the Best Tree

. . .

Now that we've obtained our performance metrics and ranked our hyperparameter "combinations", we can 

1. finalize our workflow with the best parameter choices, and
2. fit that model to our training data

. . . 

::::{.columns}

:::{.column width="50%"}

```{r}
#| echo: true
#| eval: false

best_tree <- dt_tune_results %>%
  select_best(metric = "rmse")

dt_wf_final <- dt_wf %>%
  finalize_workflow(best_tree)

dt_fit <- dt_wf_final %>%
  fit(ames_train)

dt_fit
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
#| eval: true

best_tree <- dt_tune_results %>%
  select_best(metric = "rmse")

dt_wf_final <- dt_wf %>%
  finalize_workflow(best_tree)

dt_fit <- dt_wf_final %>%
  fit(ames_train)

dt_fit
```

:::

::::

## Visualizing the Decision Tree

. . . 

A major strength of decision trees is how interpretable and intuitive they are (if we can see them)

. . . 

```{r}
#| echo: true
#| eval: true

library(rpart.plot)

dt_fit_for_plot <- dt_fit %>%
  extract_fit_engine()

rpart.plot(dt_fit_for_plot, roundint = FALSE)
```

## Let Us Now Crank Up the Tunes!!!

<center>

<iframe src="https://giphy.com/embed/elQ5RkAm90DZpirhRg" width="480" height="480" style="" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/windows-windows95-windows95man-elQ5RkAm90DZpirhRg">via GIPHY</a></p>

</center>

## Tuning Over a Workflow Set

. . . 

So we've got a decision tree model optimized over the tree depth

. . . 

All we know is that this is the optimal *decision tree*

. . . 

What about those other models we talked about??

+ A LASSO?
+ A Random Forest?
+ A Gradient Boosting Ensemble?

. . . 

Let's tune all of these models!

## Creating the Model Specifications

. . . 

```{r}
#| echo: true
#| eval: true

lasso_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

rf_spec <- rand_forest(mtry = tune(), trees = tune()) %>%
  set_engine("ranger") %>%
  set_mode("regression")

xgb_spec <- boost_tree(mtry = tune(), trees = tune(), learn_rate = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

## A Recipe and Workflow Set

. . . 

Let's use the same *recipe* for all of the models

. . . 

```{r}
#| echo: true
#| eval: true

rec <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_impute_knn(all_predictors()) %>%
  step_other(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())
```

. . . 

Now let's create the recipe and model lists

```{r}
#| echo: true
#| eval: true

rec_list <- list(rec = rec)
model_list <- list(lasso = lasso_spec, rf = rf_spec, xgb = xgb_spec)
```

. . . 

And package the lists together into a workflow set

. . .

```{r}
#| echo: true
#| eval: true

model_wfs <- workflow_set(rec_list, model_list, cross = TRUE)
```

## Constructing a Hyperparameter Grid

. . . 

We need to define our *hyperparameter grid* which declares the combinations of hyperparameters we will tune over

. . . 

::::{.columns}

:::{.column width="50%"}

```{r}
#| echo: true
#| eval: true

param_grid <- grid_regular(
  penalty(),
  mtry(range = c(1, 10)),
  trees(),
  learn_rate(),
  levels = 4
)
```

:::

:::{.column width="50%"}

Note that we've just defined a regular grid of $4^4 = 256$ hyperparameter combinations

:::

::::

. . . 

We're cross-validating our models over five folds and we have three model classes being fit, so now counting model classes, grid combinations, and folds, we have $3\times 5\times 256 = 3840$ model fits to run

. . . 

This total ignores that our ensembles consist of 50 - 2000 models themselves

. . . 

<center>

<iframe src="https://giphy.com/embed/l3q2Bz3mZjEAl4VEY" width="480" height="266" style="" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/nyfw-nyfw-feb-2017-l3q2Bz3mZjEAl4VEY">via GIPHY</a></p>

</center>

## Tuning the Models in the Workflow Set

. . .

We'll start by defining the grid controls

. . . 

```{r}
#| echo: true
#| eval: true

grid_ctrl <- control_grid(
  save_pred = TRUE,
  parallel_over = "everything",
  save_workflow = TRUE
)
```

. . . 

Using *parallel processing* will help reduce the time it takes to tune all of our models over all the hyperparameter combinations -- let's tune!

. . .

```{r}
#| echo: true
#| eval: true

n_cores <- parallel::detectCores()
cluster <- parallel::makeCluster(n_cores - 1, type = "PSOCK")
doParallel::registerDoParallel(cluster)

tictoc::tic() #Time the tuning procedure: start

wfs_tune_results <- model_wfs %>%
  workflow_map(
    seed = 123,
    resamples = ames_folds,
    grid = 10,
    control = grid_ctrl
  )

tictoc::toc() #Time the tuning procedure: end
```

## Visualizing Results

```{r}
#| echo: true
#| eval: true
#| fig-align: center

wfs_tune_results %>%
  autoplot()
```

## Our Best Model

. . .

It looks like the *gradient boosting ensemble* produced our best results in terms of both RMSE and $R^2$

. . .

Let's verify this for RMSE

. . .

```{r}
#| echo: true
#| eval: false

wfs_tune_results %>% 
  rank_results() %>%
  filter(.metric == "rmse")
```

```{r}
#| echo: false
#| eval: true

wfs_tune_results %>% 
  rank_results() %>%
  head(n = 20) %>%
  filter(.metric == "rmse") %>%
  kable() %>%
  kable_styling()
```

## Seeing Best Hyperparameter Combinations

```{r}
#| echo: true
#| eval: true
#| fig-align: center
#| fig-width: 12

wfs_tune_results %>%
  autoplot(metric = "rmse", id = "rec_xgb")
```

## Extracting Best Hyperparameter Combinations

. . .

Let's extract those best hyperparameter settings

. . .

```{r}
#| echo: true
#| eval: false

best_model <- wfs_tune_results %>%
  rank_results("rmse", select_best = TRUE) %>%
  head(1) %>%
  pull(wflow_id)

best_wf <- wfs_tune_results %>%
  extract_workflow_set_result(best_model)
  
best_params <- best_wf %>%
  select_best(metric = "rmse")

best_params
```

```{r}
#| echo: false
#| eval: true

best_model <- wfs_tune_results %>%
  rank_results("rmse", select_best = TRUE) %>%
  head(1) %>%
  pull(wflow_id)

best_wf <- wfs_tune_results %>%
  extract_workflow_set_result(best_model)
  
best_params <- best_wf %>%
  select_best(metric = "rmse")

best_params %>%
  kable() %>%
  kable_styling()
```

## Let's Fit this Best Model

. . . 

We know our best-performing model was a *gradient boosting ensemble*, and we've stored our optimized hyperparameters in `best_params`

. . .

Let's construct and fit that model!

. . .

```{r}
#| echo: true
#| eval: true

best_params <- best_params %>%
  as.list()

set.seed(123)

xgb_spec <- boost_tree(
  trees = best_params$trees,
  mtry = best_params$mtry,
  learn_rate = best_params$learn_rate
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")

xgb_wf <- workflow() %>%
  add_model(xgb_spec) %>%
  add_recipe(rec)

xgb_fit <- xgb_wf %>%
  fit(ames_train)
```

## Extracting Important Predictors

. . . 

Ensembles are not generally known for their interpretability

. . .

We can, however, identify the predictors which are most important to the ensemble using the `vip()` function from the `{vip}` package

. . .

::::{.columns}

:::{.column width="15%"}

:::

:::{.column width="70%}

```{r}
#| echo: true
#| eval: true

library(vip)

xgb_fit %>%
  extract_fit_engine() %>%
  vip()
```

:::

:::{.column width="15%"}

:::

::::

## Assessing Performance

. . .

Let's assess our model's performance on our test data

. . .

::::{.columns}

:::{.column width="50%"}

```{r}
#| echo: true
#| eval: true

xgb_results <- xgb_fit %>%
  augment(ames_test) %>%
  select(Sale_Price, .pred)

xgb_results %>%
  ggplot() + 
  geom_point(aes(x = Sale_Price, y = .pred),
             alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", alpha = 0.75) +
  labs(
    title = "Predictions versus Actuals",
    x = "Sale Price",
    y = "Prediction"
  ) + 
  coord_equal()
```

:::

:::{.column width="50%"}

```{r}
#| echo: true
#| eval: false

xgb_results %>%
  rmse(Sale_Price, .pred)
```

```{r}
#| echo: false
#| eval: true

xgb_results %>%
  rmse(Sale_Price, .pred) %>%
  kable() %>%
  kable_styling()
```

:::

::::

## Recap

+ We can use cross-validation to *tune hyperparameters*
  
  + This means that, after tuning, we can justify our hyperparameter choices

+ We can tune over a *workflow set* which contains combinations of model specifications and recipes

  + This allows us to not only find an optimal model from a particular class, but an optimal model (including class and hyperpameter settings) from a collection
  
+ Tuning allows you to improve "off the shelf" models, often by a significant margin

+ Our best untuned model to predict `Sale_Price` was a *Random Forest* that had a cross-validation RMSE of \$28,606.09, while this *tuned gradient boosting ensemble* had a cross-validation RMSE of \$24,224.78

. . . 

::::{.columns}

:::{.column width="50%"}

<center>

<iframe src="https://giphy.com/embed/7yojoQtevjOCI" width="480" height="346" style="" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/profile-notoverthehill-tomdds-7yojoQtevjOCI">via GIPHY</a></p>

</center>

:::

:::{.column width="50%"}

<center>

<br/>
<br/>
<font size="120pt">Go forth and model responsibly!</font>

</center>

:::

::::

## What's Next??? 

. . . 

**Next Time:** Thanksgiving In-Class Modeling Competition (there will be prizes, again!)

. . . 

**Then:** Hyperparameters, tuning, and other regressors workshop (and help with HW)

. . . 

**After That:** Final projects for the remainder of the semester

## Stop Parallelization

```{r}
#| echo: true
#| eval: true

parallel::stopCluster(cluster)
```