---
title: "Overview of Statistical Learning (and Competition Overview)"
author: Dr. Gilbert
format: revealjs
fontsize: 24pt
date: today
date-format: long
theme: serif
incremental: true
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)

theme_set(theme_bw(base_size = 20))
```

## Statistical Learning in Pictures

. . . 

```{r}
#| echo: false
#| message: false

num_pts <- 15
set.seed(123)
x <- runif(num_pts, 0, 100)
y <- (x - 75)^2 + rnorm(num_pts, 0, 350)

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  labs(title = "Our observed data",
       x = "x", y = "y")
  
```

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

num_pts <- 15
set.seed(123)
x <- runif(num_pts, 0, 100)
y <- (x - 75)^2 + rnorm(num_pts, 0, 350)

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  labs(title = "Our observed data",
       x = "x", y = "y")
  
```

+ **Goal:** Build a model $\displaystyle{\mathbb{E}\left[y\right] = \beta_0 + \beta_1 x}$ to predict $y$, given $x$.
+ **Generalized Goal:** Build a model $\displaystyle{\mathbb{E}\left[y\right] = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \cdots + \beta_k x_k}$ to predict $y$ given features $x_1, \cdots, x_k$.

  + $\beta_i$'s are *parameters*, *learned* from training data.

## Statistical Learning in Pictures

. . .

```{r}
#| echo: false
#| message: false

x_vals <- seq(0, 100, length.out = 500)
y_vals <- -750 + 50*x_vals

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals),
            color = "purple",
            lwd = 2,
            alpha = 0.75) + 
  labs(title = 'How "good" is the purple model?',
       x = "x", y = "y")
```

+ This model doesn't capture the general trend between our observed $x$ and $y$ pairs.

## Statistical Learning in Pictures

. . . 

```{r}
#| echo: false
#| message: false

x_vals <- seq(0, 100, length.out = 500)
y_vals <- 7000 - 60*x_vals

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals),
            color = "purple",
            lwd = 2,
            alpha = 0.75) + 
  labs(title = 'Okay, how about this one?',
       x = "x", y = "y")
```

+ Better job capturing the general trend (*sort of*).
+ Larger $x$ values are associated with smaller $y$ values.

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

preds <- 7000 - 60*x

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "Check out those prediction errors",
       x = "x", y = "y")
```

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

preds <- 7000 - 60*x

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_segment(aes(x = x, xend = x,
                   y = y, yend = preds),
               color = "red", 
               size = 3,
               alpha = 0.5) + 
  labs(title = "Check out those prediction errors",
       x = "x", y = "y",
       caption = "Reducible error remains")
```

+ Always predicting too high!

  + We should overpredict sometimes and underpredict others. The *average* error should be $0$.

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

my_data <- tibble(x = x, y = y)

lin_reg_spec <- linear_reg()
lin_reg_rec <- recipe(y ~ x, data = my_data)
lin_reg_wf <- workflow() %>%
  add_model(lin_reg_spec) %>%
  add_recipe(lin_reg_rec)
lin_reg_fit <- lin_reg_wf %>%
  fit(my_data)

coefs <- lin_reg_fit %>%
  extract_fit_engine() %>%
  tidy() %>%
  pull(estimate)

y_vals_lin <- coefs[1] + x_vals*coefs[2]

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "Getting better?",
       x = "x", y = "y")
```

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

preds <- coefs[1] + x*coefs[2]

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_segment(aes(x = x, xend = x,
                   y = y, yend = preds),
               color = "red", 
               size = 3,
               alpha = 0.5) + 
  labs(title = "Getting better?",
       x = "x", y = "y",
       caption = "Remaining error is irreducible (at least with this model form)")
```

## Statistical Learning in Pictures

```{r}
#| echo: false
#| message: false

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_segment(aes(x = x, xend = x,
                   y = y, yend = preds),
               color = "red", 
               size = 3,
               alpha = 0.5) + 
  labs(title = "Getting better?",
       x = "x", y = "y")
```

+ Capturing the general trend?

  + ...*mostly*

+ Balanced errors?

  + ✓

## How it works

. . . 

In this case, we have $\mathbb{E}\left[y\right] = \beta_0 + \beta_1\cdot x$ and we find $\beta_0$ (intercept) and $\beta_1$ (slope) to minimize the quantity

. . . 

$$\sum_{i = 1}^{n}{\left(y_{\text{obs}_i} - y_{\text{pred}_i}\right)^2}$$

## How it works

In this case, we have $\mathbb{E}\left[y\right] = \beta_0 + \beta_1\cdot x$ and we find $\beta_0$ (intercept) and $\beta_1$ (slope) to minimize the quantity

$$\sum_{i = 1}^{n}{\left(y_{\text{obs}_i} - \left(\beta_0 + \beta_1\cdot x_{\text{obs}_i}\right)\right)^2}$$

+ Changing $\beta_0$ and/or $\beta_1$ will change this sum.

## How it works

```{r}
#| echo: false
#| message: false

resids <- y - preds

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_rect(aes(xmin = x, xmax = ifelse(x > 50, x - resids*(100/15000), x + resids*(100/15000)),
                ymin = preds, ymax = preds + resids),
               color = "red",
               fill = "red",
               alpha = 0.25) + 
  scale_x_continuous(labels = NULL) + 
  scale_y_continuous(labels = NULL) + 
  labs(title = "Residual Sum of Squares",
       x = "x", y = "y",
       caption = "Note: Not to scale -- axis marks removed for visual purposes")
```

## Interpreting the Model

. . . 

::::{.columns}

:::{.column width="50%"}

```{r}
ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "What does the model tell us?",
       x = "x", y = "y")
```

:::

:::{.column width="50%}

$\displaystyle{\mathbb{E}\left[y\right] = 3783.21 - 45.3\cdot x}$

:::

::::

+ The expected value of $y$ when $x = 0$ is $3783.21$.
+ As $x$ increases, we expect $y$ to decrease (on average).
+ Given a unit increase in $x$, we expect $y$ to decrease by about $45.3$ units.

## Interpreting the Model

::::{.columns}

:::{.column width="50%"}

```{r}
ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "What does the model tell us?",
       x = "x", y = "y")
```

:::

:::{.column width="50%}

$\displaystyle{\mathbb{E}\left[y\right] = 3783.21 - 45.3\cdot x}$

:::

::::

:::{.nonincremental}

+ The <font color="purple">expected</font> value of $y$ when $x = 0$ is $3783.21$.
+ As $x$ increases, we <font color="purple">expect</font> $y$ to decrease (on average).
+ Given a unit increase in $x$, we <font color="purple">expect</font> $y$ to decrease by <font color="purple">about</font> $45.3$ units.

:::

. . .

**Approach to Model Interpretation:** In general, we'll interpret the intercept (when appropriate) and the <font color="purple">expected</font> effect of a unit change in each predictor on the response

## Can we find a *better* model? 

. . . 

```{r}
#| echo: false
#| message: false

my_data <- tibble(x = x, y = y)

poly_reg_spec <- linear_reg()
poly_reg_rec <- recipe(y ~ x, data = my_data) %>%
  step_poly(x, degree = 7)
poly_reg_wf <- workflow() %>%
  add_model(poly_reg_spec) %>%
  add_recipe(poly_reg_rec)
poly_reg_fit <- poly_reg_wf %>%
  fit(my_data)

y_vals_poly <- poly_reg_fit %>%
  predict(tibble(x = x_vals)) %>%
  pull(.pred)

coefs <- poly_reg_fit %>%
  extract_fit_engine() %>%
  tidy() %>%
  pull(estimate)

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_poly),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "Wowzers!",
       x = "x", y = "y")
```

## Can we interpret this model?

```{r}
#| echo: false
#| message: false

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_poly),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "What a great model!",
       x = "x", y = "y")
```

+ The equation is $\mathbb{E}\left[y\right] \approx 1202 - 4911x +3156x^2 + 784x^3 +\\ 409x^4 -215x^5 -7x^6 -516x^7$
+ No thanks...

## Do we expect this model to generalize well?

```{r}
#| echo: false
#| message: false

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_poly),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "Great fit on past data",
       x = "x", y = "y")
```

## Do we expect this model to generalize well?

```{r}
#| echo: false
#| message: false

num_new <- 10
set.seed(456)
x_new <- c(runif(num_new, 0, 100), 100, 0)
y_new <- (x_new - 75)^2 + rnorm(num_new + 2, 0, 350)

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_poly),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_point(aes(x = x_new, y = y_new),
             color = "red",
             size = 3) +
  labs(title = "Oh no! Bigger (unexpectedly large) errors with new data",
       x = "x", y = "y")
```

+ Especially near $x = 0$ and $x = 100$

## Is there a happy medium?

. . . 

```{r}
#| echo: false
#| message: false

quad_reg_spec <- linear_reg()
quad_reg_rec <- recipe(y ~ x, data = my_data) %>%
  step_poly(x, degree = 2)
quad_reg_wf <- workflow() %>%
  add_model(quad_reg_spec) %>%
  add_recipe(quad_reg_rec)
quad_reg_fit <- quad_reg_wf %>%
  fit(my_data)

y_vals_quad <- quad_reg_fit %>%
  predict(tibble(x = x_vals)) %>%
  pull(.pred)

coefs <- quad_reg_fit %>%
  extract_fit_engine() %>%
  tidy() %>%
  pull(estimate)

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_quad),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  labs(title = "Something less wiggly",
       x = "x", y = "y")
```

## Is there a happy medium?

```{r}
#| echo: false
#| message: false

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) +
  geom_line(aes(x = x_vals, 
                y = y_vals_quad),
            color = "purple",
            lwd = 2,
            alpha = 0.75) +
  geom_point(aes(x = x_new, y = y_new),
             color = "red",
             size = 3) + 
  labs(title = "Something less wiggly",
       x = "x", y = "y")
```

+ Fits old and new observations similarly well
+ Equation $\displaystyle{\mathbb{E}\left[y\right] \approx 1202 -4912x + 3156x^2}$

  + We'll be able to interpret this
  
## How do we know what model is *right*?

. . . 

::::{.columns}

:::{.column width="50%"}

```{r}
#| fig.height: 8

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) + 
  geom_line(aes(x = x_vals, y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.9) + 
  geom_line(aes(x = x_vals, y = y_vals_quad),
            color = "darkgreen",
            lwd = 2,
            alpha = 0.9) + 
  geom_line(aes(x = x_vals, y = y_vals_poly),
            color = "orange",
            lwd = 2,
            alpha = 0.9) + 
  geom_point(aes(x = x_new, y = y_new),
             color = "red",
             size = 3) +
  labs(title = "All three models",
       x = "x",
       y = "y")
```

:::

:::{.column width="50%"}

+ The <font color="purple">purple</font> model is *too straight*
+ The <font color="orange">orange</font> model is *too wiggly*
+ The <font color="green">green</font> model is *just right*

:::

::::

## How do we know what model is *right*?

. . . 

::::{.columns}

:::{.column width="50%"}

```{r}
#| fig.height: 8

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3) + 
  geom_line(aes(x = x_vals, y = y_vals_lin),
            color = "purple",
            lwd = 2,
            alpha = 0.3) + 
  geom_line(aes(x = x_vals, y = y_vals_quad),
            color = "darkgreen",
            lwd = 2) + 
  geom_line(aes(x = x_vals, y = y_vals_poly),
            color = "orange",
            lwd = 2,
            alpha = 0.3) + 
  geom_point(aes(x = x_new, y = y_new),
             color = "red",
             size = 3) +
  labs(title = "All three models",
       x = "x",
       y = "y")
```

:::

:::{.column width="50%"}

+ We don't want to wait for new data to know we are wrong.

  + Use some of our available data for training
  + And the rest for validation

:::

::::

## Okay, but our predictions are all wrong...literally!

::::{.columns}

:::{.column width="50%"}

```{r}
#| fig.height: 8

ggplot() + 
  geom_point(aes(x = x, y = y),
             size = 3,
             alpha = 0.25) + 
  geom_line(aes(x = x_vals, y = y_vals_quad),
            color = "darkgreen",
            lwd = 2) + 
  geom_point(aes(x = x_new, y = y_new),
             color = "red",
             size = 3) +
  labs(title = "Our best model",
       x = "x",
       y = "y")
```

:::

:::{.column width="50%"}

+ *All models are wrong, but some are useful*, George Box (1976)
+ Predictions will be wrong but, with some assumptions, they have value

:::

::::

## Necessary Assumptions

. . .

For model and predictions

+ Training data are random and representative of population.

  + Otherwise, we should not be modeling this way.

+ Residuals (prediction errors) are normally distributed with mean $\mu = 0$ and constant standard deviation $\sigma$.

  + Allows construction of confidence intervals around predictions (making our models *right*).

## Necessary Assumptions

. . .

For interpretations of coefficients (statistical learning / inference)

+ No multicollinearity (predictors aren't correlated with one another)
  
## Summary

+ Building models $\mathbb{E}\left[y\right] = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \cdots + \beta_k x_k$
+ Predicting a *numerical* response ($y$) given features ($x_i$)
+ Need data to build the models -- some for training, some for validation

  + The $\beta_i$'s are *parameters* whose values are *learned/estimated* from *training* data
+ Model predictions will be wrong
+ As long as standard deviation of residuals (prediction errors) is constant, we can build meaningful confidence intervals for predictions
+ Can interpret models to gain insight into relationships between predictor(s) and response

## Competition Information

+ Predict *Door Dash* delivery times

  + Time from placing order to delivery
  
+ **Response:** delivery time
+ **Predictors:** `market_id`, `order_time`, `delivery_time`, `store_id`, `cuisine_type`, `order_protocol`, `items_in_order`, `subtotal_cost`, `distinct_items_in_order`, `min_item_price`, `max_item_price`, `dashers_working`, `busy_dashers`, `outstanding_orders`,  `model_1_estimate`, `model_2_estimate`

## How this works

+ Six assignments to guide you.
+ You'll take data provided by *Door Dash* and build models.
+ Kaggle will assess the predictions from everyone's models.
+ Live leaderboard, using only part of the competition data so you know approximately where you stand.
+ You'll talk with eachother about why/how you have different scores. 
+ People *might* share their strategies, or they *might not* -- this is a competition after all.

## Why?

+ Interest in homework assignments generally ends after they're turned in and graded.
+ Competition assignments and environment ask you to iterate on previous work.
+ You'll almost surely be interested in what other people have done, especially if their models have performed better than yours.
+ You'll talk with eachother about strategies and modeling choices.
+ You'll be motivated to improve your model even between assignments.

## What past students say

+ The competition is fun
+ It is motivating
+ I learned more because I wanted to place better in the competition
+ Talking with others about their models made me more confident in my understanding of course material

## What you are building

+ An analytics report
+ You'll be building models *and* (more importantly) writing about your modeling choices and the performance of your models
+ Six assignments -- each focusing on part(s) of the modeling process and analytics report
+ Prepares you for the final project, where you'll do all this over again on a data set you identified and care about

## Next Time...

+ Way fewer slides! 🤕
+ An introduction to R
+ Getting our hands dirty!

. . . 

**Homework:** Start Competition Assignment 1 -- join the competition, read the details, download the data, and start writing a *Statement of Purpose*