---
title: "Welcome to R; A Quick Overview"
author: Dr. Gilbert
format: 
  revealjs:
    smaller: true
date: today
date-format: long
theme: serif
incremental: true
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)

theme_set(theme_bw(base_size = 20))
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

## The Highlights

+ Interacting with R
+ The R dialects
+ Installing and loading packages
+ Reading in data
+ First interactions with data in R
+ The pipe operator (`%>%` or `|>`)
+ Manipulating and transforming data with R
+ All changes are temporary unless you make them permanent

## Interacting with R

+ We'll interact with R via RStudio in MAT300.

  + In the future, you can use R with other IDEs, like VS code.
  
+ We'll be using Quarto Documents/Notebooks for everything we do.

  + Notebooks allow us to fully document and share analyses even with non-technical people.
  
## Let's Get Started

1. Open RStudio
2. Create a new *project* by navigating to `File -> New Project`

    + Choose to create the project in a new directory (folder)
    + Name this something like `MAT300` -- you'll include all of your notebooks for our class in this project
    + **Note:** If you'd like to use GitHub and manage your project space using a repository, come see me in *office hours* or reach out on *Slack* and I'll help you get set up.
  
3. Now that you are in your new project space, create a new Quarto Document by navigating to `File -> New File -> Quarto Document`

    + You can edit the fields, or accept the defaults -- it's up to you
  
## The R Dialects

There are three main dialects in R

+ Base-R
+ `data.table` (speed)
+ Tidy-R / `tidyverse` (readability and consistency)

  + We use this one. When you search for help -- include "*tidyverse*" in your query.
  
. . .

> **Note:** *R dialects* just refers to how we choose to write R code and which functions we prioritize -- dialects can be (and often are) mixed.

## Installing and Loading Packages

+ `install.packages("PACKAGE_NAME")` to install a package

  + You only do this once
  + Do it in the *console*

+ `library(PACKAGE_NAME)` to load the package

  + Do this near the top of your Quarto Document, in a *setup* chunk
  
. . .

$\bigstar$ Install the `{tidyverse}` and load it in your Quarto Notebook

## Reading in Data

. . .

**ASIDE:** We store objects in R with the arrow operator (`<-`)

. . .

```{r}
#| echo: true
#| eval: false

x <- 2
```

. . .

**Reading Data:** To read data, we use a function of the form `read_*()`.

. . .

```{r}
#| echo: true
#| eval: false

data <- read_csv("PATH_TO_CSV_FILE")
```

+ Requires the `{tidyverse}` (or at least `{readr}`) to be loaded
+ Similar functions exist for reading other file formats
+ Some require other packages (`{readxl}` or `{haven}` are common)

. . .

$\bigstar$ Read the `ames` dataset into your Quarto Notebook [from this link](https://raw.githubusercontent.com/koalaverse/homlr/master/data/ames.csv)

. . .

I'll post the link in Slack
  
## First Interactions with Data

+ `head()` to view first six rows
+ `glimpse()` to view dimensions and data types
+ `skim()` from `{skimr}` for much more detail

. . . 

$\bigstar$ Try these functions on your `ames` data

## The Pipe Operator 

+ Pipes (`%>%` or `|>`) make code more readable and allow chaining of functions together
+ Object to the left of the pipe becomes first argument to the function after the pipe
+ Read the pipe to mean "*and then*"

. . .

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|1|2|"

ames %>%
  head()
```

## Manipulating and Transforming Data

+ `filter()` to return only desired records
+ `select()` to return only desired columns
+ `summarize()` to compute summaries on a table
+ `groupby()` to create groups in a table
+ `mutate()` to create new columns or change existing ones

. . .

$\bigstar$ How might we use these functions? Let's ask some questions and then try to answer them using the functions above.

. . . 

Add your work to your Quarto Notebook

## All Changes are Temporary

```{r}
#| echo: true
#| eval: false

penguins %>%
  filter(species == "Gentoo") %>%
  groupby(island) %>%
  summarize(
    avg_mass = mean(body_mass_g)
  )
```

+ Start with the `penguins` data frame, *and then*
+ filter to just the Gentoo species, *and then*
+ group by island, *and then*
+ calculate average penguin body mass for each group

**Note:** `penguins` data frame is not permanently altered here

## Until We Make Them Permanent

. . .

```{r}
#| echo: true
#| eval: false

penguins <- penguins %>%
  filter(species == "Gentoo") %>%
  groupby(island) %>%
  summarize(
    avg_mass = mean(body_mass_g)
  )
```

. . .

Now the change is permanent because we've stored the result

+ Notice the use of the arrow operator (`<-`)
+ Be careful overwriting existing objects -- think about whether you:
  
  + might need the old object again
  + would be better off creating a new object (variable)
  + even need to store the result at all
    
## Let's Practice

. . .

**Reminder:** You have a fully complete (and documented) notebook using the `mpg` data on the course webpage

. . .

Use this time to continue playing with the `ames` housing data

1. Write down and answer some interesting questions that might use functionality discussed in this slide deck -- start simple and then build up to questions that might be more complex
2. Try documenting your work by including text descriptions in addition to the code chunks
3. Save your QMD file
4. Use the blue *render* button to convert your markdown document into a beautiful HTML document and enjoy the fruits of your labor!

    + 🤔 Ponder an existence where you never need to open MS Word again! 🤔

. . .

Don't worry if your document looks quite plain for now, we'll have a full class meeting devoted to using markdown syntax in Quarto effectively