---
title: "Welcome to R; A Quick Overview"
author: Dr. Gilbert
format: 
  revealjs:
    smaller: true
date: today
date-format: long
theme: serif
incremental: true
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(countdown)

theme_set(theme_bw(base_size = 20))
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}

a:link{
  color: purple;
}

a:visited{
  color: purple;
}
```

## The Highlights

+ Interacting with R
+ The R dialects
+ Installing and loading packages
+ Reading in data
+ First interactions with data in R
+ The pipe operator (`%>%` or `|>`)
+ Manipulating and transforming data with R
+ All changes are temporary unless you make them permanent

## Interacting with R

+ We'll interact with R via RStudio in MAT300.

  + In the future, you can use R with other IDEs, like VScode or *Positron*.
  
+ We'll be using Quarto Documents/Notebooks for everything we do.

  + Notebooks allow us to fully document and share analyses even with non-technical people.
  
## Let's Get Started

1. Open RStudio
2. Create a new *project* by navigating to `File -> New Project`

    + Choose to create the project in a new directory (folder)
    + Name this something like `MAT300` -- you'll include all of your notebooks for our class in this project
    + **Note:** If you'd like to use GitHub and manage your project space using a repository, come see me in *office hours* or reach out on *Slack* and I'll help you get set up.
  
3. Now that you are in your new project space, create a new Quarto Document by navigating to `File -> New File -> Quarto Document`

    + You can edit the fields, or accept the defaults -- it's up to you
  
## The R Dialects

There are three main dialects in R

+ Base-R
+ `data.table` (speed)
+ Tidy-R / `tidyverse` (readability and consistency)

  + We use this one. When you search for help -- include "*tidyverse*" in your query.
  
. . .

> **Note:** *R dialects* just refers to how we choose to write R code and which functions we prioritize -- dialects can be (and often are) mixed.

## Installing and Loading Packages

+ `install.packages("PACKAGE_NAME")` to install a package

  + You only do this once
  + Do it in the *console*

+ `library(PACKAGE_NAME)` to load the package

  + Do this near the top of your Quarto Document, in a *setup* chunk
  
. . .

$\bigstar$ Install the `{tidyverse}` and load it in your Quarto Notebook. Do the same with `{tidymodels}`.

## Reading in Data

. . .

**ASIDE:** We store objects in R with the arrow operator (`<-`)

. . .

```{r}
#| echo: true
#| eval: false

x <- 2
```

. . .

**Reading Data:** To read data, we use a function of the form `read_*()`.

. . .

```{r}
#| echo: true
#| eval: false

data <- read_csv("PATH_TO_CSV_FILE")
```

+ Requires the `{tidyverse}` (or at least `{readr}`) to be loaded
+ Similar functions exist for reading other file formats
+ Some require other packages (`{readxl}` or `{haven}` are common)

. . .

$\bigstar$ Read this `airbnb` dataset into your Quarto Notebook [from this link](https://raw.githubusercontent.com/agmath/agmath.github.io/master/data/AirBnB.csv)

. . .

I'll post the link in Slack

. . . 

> **Note:** This AirBnB Europe data was uploaded to Kaggle by Dipesh Khemani. The [original dataset can be found here](https://www.kaggle.com/datasets/dipeshkhemani/airbnb-cleaned-europe-dataset?select=Aemf1.csv).
  
## First Interactions with Data

+ `head()` to view first six rows
+ `glimpse()` to view dimensions and data types
+ `skim()` from `{skimr}` for much more detail

. . . 

$\bigstar$ Try these functions on your `airbnb` data

## The Pipe Operator 

+ Pipes (`%>%` or `|>`) make code more readable and allow chaining of functions together
+ Object to the left of the pipe becomes first argument to the function after the pipe
+ Read the pipe to mean "*and then*"

. . .

```{r}
#| echo: true
#| eval: false
#| code-line-numbers: "|1|2|"

penguins %>%
  head()
```

. . . 

$\bigstar$ Rewrite the functions you used to explore your data with *pipes*

## Manipulating and Transforming Data

+ `filter()` to return only desired records
+ `select()` to return only desired columns
+ `summarize()` to compute summaries on a table
+ `group_by()` to create groups in a table
+ `mutate()` to create new columns or change existing ones

. . .

$\bigstar$ How might we use these functions? Write down some questions that could be answered using the functions described above. Start with a couple very simple questions and then work up to questions whose answers might be more complex to find.

## Manipulating and Transforming Data

:::{.nonincremental}

+ `filter()` to return only desired records
+ `select()` to return only desired columns
+ `summarize()` to compute summaries on a table
+ `group_by()` to create groups in a table
+ `mutate()` to create new columns or change existing ones

:::

$\bigstar$ How might we use these functions? Write down some questions that could be answered using the functions described above. Start with a couple very simple questions and then work up to questions whose answers might be more complex to find.

$\bigstar$ We'll try answering some of those questions now!

## All Changes are Temporary

```{r}
#| echo: true
#| eval: false

penguins %>%
  filter(species == "Gentoo") %>%
  group_by(island) %>%
  summarize(
    avg_mass = mean(body_mass_g)
  )
```

+ Start with the `penguins` data frame, *and then*
+ filter to just the Gentoo species, *and then*
+ group by island, *and then*
+ calculate average penguin body mass for each group

**Note:** `penguins` data frame is not permanently altered here

## Until We Make Them Permanent

. . .

```{r}
#| echo: true
#| eval: false

penguins <- penguins %>%
  filter(species == "Gentoo") %>%
  group_by(island) %>%
  summarize(
    avg_mass = mean(body_mass_g)
  )
```

. . .

Now the change is permanent because we've stored the result

+ Notice the use of the arrow operator (`<-`)
+ Be careful overwriting existing objects -- think about whether you:
  
  + might need the old object again
  + would be better off creating a new object (variable)
  + even need to store the result at all
    
## Let's Practice

. . .

**Reminder:** You have a fully complete (and documented) notebook using car data (`mpg`) on the course webpage -- note this data is different than the `airbnb` data you are working with today

. . .

Use this time to continue playing with the `airbnb` pricing data.

1. Save your QMD file
2. Use the blue *render* button to convert your markdown document into a beautiful HTML document and enjoy the fruits of your labor!

    + 🤔 Ponder an existence where you never need to open MS Word again! 🤔
    
3. Write down and answer additional interesting questions that might use functionality discussed in this slide deck -- start simple and then build up to questions that might be more complex
4. Document your work by including text descriptions alongside the code chunks

. . .

Don't worry if your document looks quite plain for now, we'll have a full class meeting devoted to using markdown syntax in Quarto effectively.

## Next Time...

<center>

<font size="120pt"><br/>Data Visualization</font>

</center>