---
title: "Data Dirty Work: Wrangling"
author: Dr. Gilbert
format: 
  revealjs:
    smaller: true
date: today
date-format: long
theme: serif
incremental: true
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(patchwork)
library(kableExtra)
library(randomNames)

options(kable_styling_bootstrap_options = c("hover", "striped"))

theme_set(theme_bw(base_size = 20))
```

```{css}
code {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

## A Quote and Setting Expectations

+ "*Life is dirty. So is your data. Get used to it.*" -- Thomas LaRock

+ Breakdown of Data Science work:

  + 80% data cleaning
  + 20% analysing, modeling, and interpreting

+ Building models is easy.

  + Formatting your data appropriately is difficult.
  + Interpreting your models takes care.

## The Highlights

+ What is *tidy* data?
+ Common scenarios prompting data cleaning / data wrangling

  + Inconsistent data entry/data collection
  + Incorrect/unexpected data types inferred
  + A single column encodes multiple variables
  + Data currently in *wide* format when we want *long* format, and vice-versa
  + Features are stored across different tables

## Tidy Data

Tabular data is *tidy* if the following three principles are satisfied.

+ Each row corresponds to a single observational unit (a record).
+ Each column corresponds to a single measured variable.
+ Each cell contains the value of the corresponding variable measured on the corresponding record.

  + **Note:** Missing values are permitted.

Simply put, data is *tidy* if it is in a format convenient for analysis.

## A Problematic Data Set {.scrollable}

```{r}
#| echo: false

set.seed(123)
sample_df <- tibble(
  id = sample(1:100, size = 6),
  month = sample(1:12, size = 6, replace = TRUE),
  day = sample(1:28, size = 6, replace = TRUE),
  year = sample(2010:2023, size = 6, replace = TRUE),
  salary = c(55000, "$58,500", "60,000", 22.75, "$1400", 72250),
  favorite_color = c("blue", "purple", "orang", "Blue", "BLUE", "orange")
)

sample_df <- sample_df %>%
  unite(date, c(month, day, year), sep = "/")

sample_df %>%
  kable() %>%
  kable_styling()

sample_df %>% 
  glimpse()
```

## A Better Version {.scrollable}

```{r}
#| echo: false

better_df <- sample_df %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(
    salary = str_replace_all(salary, "[$,]", ""),
    favorite_color = tolower(favorite_color)
  ) %>%
  mutate(
    month = as.numeric(month),
    day = as.numeric(day),
    year = as.numeric(year),
    salary = as.numeric(salary)
  ) 

better_df %>%
  kable() %>%
  kable_styling()

better_df %>%
  glimpse()
```

+ Problems still remain. What are they?

## A Dataset That Is Too Wide

```{r}
#| echo: false

set.seed(234)
sample_df <- tibble(
  name = randomNames(6),
  `2019` = round(runif(6, min = 2.1, 8.2), 1),
  `2020` = round(runif(6, min = 7.8, 13.1), 1),
  `2021` = round(runif(6, min = 4.6, 10.1), 1),
  `2022` = round(runif(6, min = 3.4, 9.3), 1),
  `2023` = round(runif(6, min = 3.4, 9.3), 1),
)

sample_df %>%
  kable() %>%
  kable_styling()
```

## A Better Version {.scrollable}

```{r}
#| echo: false

sample_df %>%
  pivot_longer(cols = `2019`:`2023`, names_to = "year", values_to = "value") %>%
  mutate(year = as.numeric(year)) %>%
  kable() %>%
  kable_styling()
```

## A Dataset That Is Too Long

```{r}
#| echo: false

sample_df <- tibble(
  name = rep(randomNames(3), 2)
) %>%
  arrange(name) %>%
  mutate(measurement = rep(c("height", "weight"), 3)) %>%
  mutate(value = rnorm(6, mean = 175, sd = 10)) %>%
  mutate(value = ifelse(measurement == "weight", value/2.4, value))

sample_df %>%
  kable() %>%
  kable_styling()
```

## A Better Version

```{r}
#| echo: false

sample_df %>%
  pivot_wider(names_from = measurement, values_from = value) %>%
  kable() %>%
  kable_styling()
```

## Data Across Different Tables {.scrollable}

```{r}
#| echo: false

set.seed(567)

items <- c("apple", "banana", "pear", "mango", "peach")

order_df <- tibble(
  name = randomNames(5),
  ordered = sample(items, 5, replace = TRUE),
  quantity = sample(1:6, 5, replace = TRUE)
)

price_df <- tibble(
  item = items,
  price = round(runif(5, 1, 5), 2)
)
```

**Orders Data Frame:**

```{r}
#| echo: false

order_df %>%
  kable() %>%
  kable_styling()
```

**Prices Data Frame:**

```{r}
#| echo: false

price_df %>%
  kable() %>%
  kable_styling()
```

## Joined Data for Analysis

```{r}
#| echo: false

order_df %>%
  left_join(price_df,
            by = c("ordered" = "item")) %>%
  kable() %>%
  kable_styling()
```

## Commonly Utilized Tools {.smaller}

+ Tools for dealing with inconsistent data entry
  + `mutate()` with `ifelse()`, `case_when()`, `str_replace()`, etc.

+ Tools for changing data types

  + `as.numeric()`, `as.integer()`, `as.factor()`, etc.
  
+ Extracting multiple variables from a single column
  
  + `separate()`

+ Moving between wide and long formats

  + `pivot_longer()` and `pivot_wider()`
  
+ Joining data from multiple tables

  + `left_join()`, `full_join()`, `inner_join()`, `anti_join()`, etc.
  + Almost always `left_join()`

