---
title: "Simple Linear Regression: Construction, Assessment, and Interpretation"
author: Dr. Gilbert
format: 
  revealjs:
    smaller: true
date: today
date-format: long
theme: serif
incremental: true
fontsize: 22pt
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(patchwork)
library(kableExtra)
tidymodels_prefer()

penguins <- palmerpenguins::penguins

penguins <- penguins %>%
  mutate(flipper_length_mm = as.numeric(flipper_length_mm),
         body_mass_g = as.numeric(body_mass_g))

set.seed(123)
penguin_splits <- initial_split(penguins)
penguins_train <- training(penguin_splits)
penguins_test <- testing(penguin_splits)

options(kable_styling_bootstrap_options = c("hover", "striped"))
options(scipen = 999)

#Set ggplot base theme
theme_set(theme_bw(base_size = 14))
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

## The Highlights

+ What is Simple Linear Regression?
+ What are we assuming?
+ Global tests for model utility and the individual term-based test.
+ Further model assessment

  + Validation metrics
  + Residual analysis
+ Model interpretation
+ Predictions

## What is Simple Linear Regression?

. . .

```{r}
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 4

p1 <- penguins_train %>%
  ggplot() +
  geom_point(aes(x = flipper_length_mm,
                 y = body_mass_g)) +
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)")

p2 <- penguins_train %>%
  ggplot() +
  geom_point(aes(x = bill_depth_mm,
                 y = body_mass_g)) +
  labs(x = "Bill Depth (mm)",
       y = "Body Mass (g)")

p1 + p2
```

. . .

**Question 1 (Inferential):** What, if anything, is the relationship between penguin flipper length and body mass?

. . . 

**Question 1 (Predictive):** Can we use penguin flipper length to predict body mass?

. . .

**Question 2 (Inferential):** What, if anything, is the relationship between penguin bill depth and body mass?

. . . 

**Question 2 (Predictive):** Can we use penguin bill depth to predict body mass?

## What is Simple Linear Regression?

```{r}
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 4

p1 + p2
```

Each of these questions can be answered by the construction and analysis of a model.

. . . 

Simple linear regression predicts a response as a linear function of a single predictor variable.

. . .

$$\mathbb{E}\left[\text{body mass}\right] = \beta_0 + \beta_1\cdot \left(\text{flipper length}\right)\\ \textbf{or}\\ \mathbb{E}\left[\text{body mass}\right] = \beta_0 + \beta_1\cdot \left(\text{bill depth}\right)$$

## What is Simple Linear Regression?

. . .

```{r}
#| message: false
#| warning: false
#| fig-align: center

lin_reg_flip_spec <- linear_reg()
lin_reg_flip_rec <- recipe(body_mass_g ~ flipper_length_mm, data = penguins_train)
lin_reg_flip_wf <- workflow() %>%
  add_recipe(lin_reg_flip_rec) %>%
  add_model(lin_reg_flip_spec)
lin_reg_flip_fit <- lin_reg_flip_wf %>%
  fit(penguins_train)

lin_reg_billd_spec <- linear_reg()
lin_reg_billd_rec <- recipe(body_mass_g ~ bill_depth_mm, data = penguins_train)
lin_reg_billd_wf <- workflow() %>%
  add_recipe(lin_reg_billd_rec) %>%
  add_model(lin_reg_billd_spec)
lin_reg_billd_fit <- lin_reg_billd_wf %>%
  fit(penguins_train)

new_data <- tibble(
  flipper_length_mm = seq(min(penguins_train$flipper_length_mm, na.rm = TRUE), max(penguins_train$flipper_length_mm, na.rm = TRUE), length.out = 250),
  bill_depth_mm = seq(min(penguins_train$bill_depth_mm, na.rm = TRUE), max(penguins_train$bill_depth_mm, na.rm = TRUE), length.out = 250)
)

new_data <- lin_reg_flip_fit %>%
  augment(new_data) %>%
  rename(flip_pred = .pred)
new_data <- lin_reg_billd_fit %>%
  augment(new_data) %>%
  rename(billd_pred = .pred)

p1 <- p1 + 
  geom_line(data = new_data,
            aes(x = flipper_length_mm,
                y = flip_pred),
            color = "blue",
            lwd = 2)

p2 <- p2 + 
  geom_line(data = new_data,
            aes(x = bill_depth_mm,
                y = billd_pred),
            color = "blue",
            lwd = 2)

p1 + p2
```

. . .

$$\mathbb{E}\left[\text{body mass}\right] = -5769 + 49.7\left(\text{flipper length}\right)\\ \textbf{or}\\ \mathbb{E}\left[\text{body mass}\right] = 7697 - 203\left(\text{bill depth}\right)$$

## Let's Play Along

. . .

<center>

$\bigstar$ As usual, I recommend that you play along during our discussion! $\bigstar$

</center>



1. Open RStudio
2. Verify that you are working within your `MAT300` project space
3. Open your most recent notebook -- the one where you built a simple linear regressor to predict the rental price of an Air BnB
4. Run all of the code chunks in that notebook
5. Describe the *inferential* and *predictive* questions you implicitly asked in pursuing the construction of our simple linear regression models from the end of our last class meeting

## What Are We Assuming?

. . .

```{r}
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 2.5
#| fig-width: 8

p1
```

. . . 

$$\mathbb{E}\left[\text{body mass}\right] = \beta_0 + \beta_1\cdot\left(\text{flipper length}\right)$$

. . . 

**Pre-Modeling Assumptions:** The conditional mean of penguin body mass is approximately a linear function of flipper length, with other factors treated as unmodeled sources of variability.

. . . 

**Post-Modeling Assumptions:** The following assumptions are made about model errors (*residuals*), to ensure that using and interpreting the model is appropriate.

+ Residuals are independent across observations.
+ Residuals have mean zero, conditional on the predictor.
+ The variance of residuals is constant across values of the predictor (homoscedasticity).
+ Residuals are approximately normally distributed (primarily to support inference).

## Global and Term-Based Model Assessments

. . .

**Global Test for Model Utility:** $\begin{array}{lcl} H_0 & : & \beta_1 = 0\\ H_a & : &\beta_1 \neq 0\end{array}$

. . .

```{r}
lin_reg_flip_fit %>%
  glance() %>%
  kable() %>%
  kable_styling()
```

. . . 

**Individual Term-Based Assessment:** $\begin{array}{lcl} H_0 & : & \beta_1 = 0\\ H_a & : &\beta_1 \neq 0\end{array}$

. . .

```{r}
lin_reg_flip_fit %>%
  extract_fit_engine() %>%
  tidy() %>%
  kable() %>%
  kable_styling()
```

. . . 

For Simple Linear Regression, the *Global Test for Model Utility* and the *Term-based test* have the same hypotheses and will have the same $p$-value. They are the same test here!

## Additional [Training] Performance Metrics

::::{.columns}

:::{.column width="45%"}

```{r}
#| eval: false
#| echo: true

lin_reg_flip_fit %>%
  glance()
```

```{r}
#| eval: true
#| echo: false

lin_reg_flip_fit %>%
  glance %>%
  pivot_longer(cols = everything(),
               names_to = "metric",
               values_to = "value") %>%
  kable() %>%
  kable_styling()
```

:::

:::{.column width="55%"}

+ $R^2_{\text{adj}} \approx 76.4\%$, so approximately 76.4% of variation in penguin body mass is explained by the linear relationship with flipper length in this model.

+ Training RMSE (`sigma`) is about 398.88, so we expect our model's typical prediction error to be about 398.88 grams.

  + Note that this estimate is likely too optimistic.

:::

::::

## Accessing and Interpreting Global and Term-Based Performance Metrics

. . . 

<center>

$\bigstar$ Let's try it! $\bigstar$ 

</center>

1. Obtain the global model performance metrics for your model

    + *Hint*. You'll need the `glance()` function for this
    
2. Analyse and interpret the result

3. Obtain the individual term-based model assessment metrics

    + *Hint*. You'll need to use `extract_fit_engine()` here

4. Analyse and interpret the result

## Additional [Validation] Performance Metrics

::::{.columns}

:::{.column width="40%"}

```{r}
#| eval: false
#| echo: true

my_metrics <- metric_set(
  rmse, rsq
  )

lin_reg_flip_fit %>%
  augment(penguins_test) %>%
  my_metrics(.pred, 
             body_mass_g)
```

```{r}
#| eval: true
#| echo: false

my_metrics <- metric_set(rmse, rsq)

lin_reg_flip_fit %>%
  augment(penguins_test) %>%
  my_metrics(.pred, body_mass_g) %>%
  select(-.estimator) %>%
  kable() %>%
  kable_styling()
```

:::

:::{.column width="60%"}

+ $R^2 \approx 74.1\%$, so approximately 74.1% of the variability in penguin body mass is accounted for by the linear relationship with flipper length when evaluated on unseen data (when taking away the training data advantage).

+ Test RMSE is about 380.82, so a more appropriate estimate for our model's typical prediction error for new data is be about 380.82 grams.

+ The test $R^2$ value is slightly worse than the corresponding training metric, while the test RMSE is an improvement over the training RMSE.

  + Generally, we should expect slightly worse performance on the test data than on training data, though individual metrics may occasionally improve due to sampling variability.

:::

::::

## Performance Metrics on Validation Data

<center>

$\bigstar$ Let's try it! $\bigstar$

</center>

1. Use the code from the previous slide to obtain model performance metrics measured on the validation data (the `test` set)
2. How do the training and validation metrics compare with one another?

## Residual Analysis

. . .

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

train_with_resids <- lin_reg_flip_fit %>%
  augment(penguins_train)

p1 <- train_with_resids %>%
  mutate(residuals = body_mass_g - .pred) %>%
  ggplot() +
  geom_histogram(aes(x = residuals, y = ..density..),
                 fill = "purple",
                 color = "black",
                 alpha = 0.75) + 
  geom_density(aes(x = residuals),
               fill = "purple",
               alpha = 0.6) + 
  labs(x = "residuals", y = "")

p2 <- train_with_resids %>%
  mutate(residuals = body_mass_g - .pred) %>%
  ggplot() +
  geom_point(aes(x = body_mass_g, y = residuals)) + 
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_smooth(aes(x = body_mass_g, y = residuals)) +
  labs(x = "Body Mass", y = "Residuals")

p3 <- train_with_resids %>%
  mutate(residuals = body_mass_g - .pred) %>%
  ggplot() +
  geom_point(aes(x = .pred, y = residuals)) + 
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_smooth(aes(x = .pred, y = residuals)) + 
  labs(x = "Predictions", y = "Residuals")

p4 <- train_with_resids %>%
  mutate(residuals = body_mass_g - .pred) %>%
  ggplot() +
  geom_point(aes(x = flipper_length_mm, y = residuals)) + 
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_smooth(aes(x = flipper_length_mm, y = residuals)) +
  labs(x = "Flipper Length", y = "Residuals")

(p1 + p2) / (p3 + p4) + plot_annotation(title = "Residual Plots")
```

. . . 

The residuals look approximately normal -- with some right skew. There does seem to be an association between the residuals and response, predictions, and flipper length though.

. . . 

Patterns in residual plots indicate that we could make a better model.

## Conducting a Residual Analysis

<center>

$\bigstar$ Let's try it! $\bigstar$

</center>

1. Attach your model's predictions to your training data
2. Compute a residuals (prediction errors) column
3. Visualize the distribution of residuals
4. Create a plot between residuals and the response
5. Create a plot between residuals and the predictions
6. Create a plot between the residuals and the predictor
7. Interpret the plots you've constructed

    + What do you notice about the plots between resduals and predictor and residuals and predictions? Why might this be?

## Model Interpretations

```{r}
#| echo: false
#| eval: true

lin_reg_flip_fit %>%
  extract_fit_engine() %>%
  tidy() %>%
  kable() %>%
  kable_styling()
```

. . .

$$\mathbb{E}\left[\text{body mass}\right] = -5768.76 + 49.67\cdot\left(\text{flipper length}\right)$$

. . .

**Interpretations:** 

+ (*Intercept*) We expect a penguin whose flipper length measures 0mm to have a mass of about -5768.76g

  + Note that this is not reasonable, and our model doesn't support this interpretation since our shortest observed flipper length was 172mm.
  + We could force the intercept to be 0, but we would observe worse fit.
  
+ (*Flipper Length*) We expect a 1mm increase in flipper length to be associated with about a 49.67g increase in penguin body mass, on average.

## Model Interpretations

<center>

$\bigstar$ Let's try it! $\bigstar$

</center>

1. Extract your model fit
2. Provide interpretations of the model coefficients

    + Is an interpretation of the intercept meaningful for your model?
    
3. Include your discussions in your Quarto Notebook

## Using the Model to Make Predictions

Consider the following questions:

1. What is the body mass of a penguin whose flipper length is 212mm?
2. What is the average body mass of *all* penguins whose flipper lengths are 212mm?

. . . 

The first question is asking about the mass of a single penguin, while the second question is asking about the average mass over all penguins with a particular characteristic.

. . .

There is more uncertainty associated with trying to answer the first question.

. . .

Our model predicts 

\begin{align} \mathbb{E}\left[\text{body mass}\right] &= -5768.76 + 49.67\left(212\right)\\
&= 4761.28\text{g}
\end{align}

as the answer to both -- we have 0% confidence in this!

## Using the Model to Make Predictions

. . .

::::{.columns}

:::{.column width="50%"}

```{r}
#| echo: true
#| eval: false

new_penguin <- tibble(
  flipper_length_mm = 212
)

lin_reg_flip_fit %>%
  predict(new_penguin)
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
#| eval: true

new_penguin <- tibble(
  flipper_length_mm = 212
)

new_penguin <- new_penguin %>%
  bind_cols(
    lin_reg_flip_fit %>%
      predict(new_penguin)
  ) %>%
  bind_cols(
    lin_reg_flip_fit %>%
      predict(new_penguin,
              type = "conf_int") %>%
      rename(.conf_lower = .pred_lower,
             .conf_upper = .pred_upper)
  ) %>%
  bind_cols(
    lin_reg_flip_fit %>%
      predict(new_penguin, type = "pred_int")
  )

lin_reg_flip_fit %>%
  predict(new_penguin) %>%
  kable() %>%
  kable_styling()
```

:::

::::

. . .

```{r}
#| message: false
#| warning: false
#| fig-align: center

point_colors <- c("training" = "black", "new" = "darkgreen")

ggplot() +
  geom_line(data = new_data,
            aes(x = flipper_length_mm, 
                y = flip_pred),
            color = "blue",
            lwd = 2,
            alpha = 0.5) + 
  geom_point(data = penguins_train,
             aes(x = flipper_length_mm,
                 y = body_mass_g,
                 color = "training"),
             alpha = 0.4) +
  geom_point(data = new_penguin,
             aes(x = flipper_length_mm,
                 y = .pred,
                 color = "new"),
             size = 6) + 
  scale_color_manual(values = point_colors) + 
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)",
       color = "Penguin Type") 

```

## Using the Model to Make Predictions

. . .

What is <font color="green">the body mass of a penguin whose flipper length is 212mm</font>?

. . . 

::::{.columns}

:::{.column width="45%"}

```{r}
#| echo: true
#| eval: false

lin_reg_flip_fit %>%
  predict(new_penguin, 
          type = "pred_int", 
          level = 0.95)
```

:::

:::{.column width="55%"}

```{r}
#| echo: false
#| eval: true

lin_reg_flip_fit %>%
  predict(new_penguin, type = "pred_int") %>%
  kable() %>%
  kable_styling()
```

:::

::::

. . . 

```{r}
#| message: false
#| warning: false
#| fig-align: center

point_colors <- c("prediction" = "darkgreen")

ggplot() +
  geom_line(data = new_data,
            aes(x = flipper_length_mm, 
                y = flip_pred),
            color = "blue",
            lwd = 2,
            alpha = 0.5) + 
  geom_point(data = penguins_train,
             aes(x = flipper_length_mm,
                 y = body_mass_g),
             color = "black",
             alpha = 0.4) +
  geom_errorbar(data = new_penguin,
                aes(x = flipper_length_mm,
                    ymin = .pred_lower,
                    ymax = .pred_upper,
                    color = "prediction"),
                alpha = 0.75,
                lwd = 1.5,
                width = 5) +
  geom_point(data = new_penguin,
             aes(x = flipper_length_mm,
                 y = .pred),
             color = "darkgreen",
             size = 6) + 
  scale_color_manual(values = point_colors) + 
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)",
       color = "Interval") 

```

## Using the Model to Make Predictions

. . .

What is <font color="purple">the *average* body mass of *all* penguins whose flipper length is 212mm</font>?

. . . 

::::{.columns}

:::{.column width="45%"}

```{r}
#| echo: true
#| eval: false

lin_reg_flip_fit %>%
  predict(new_penguin, type = "conf_int", level = 0.95)
```

:::

:::{.column width="55%"}

```{r}
#| echo: false
#| eval: true

lin_reg_flip_fit %>%
  predict(new_penguin, type = "conf_int") %>%
  kable() %>%
  kable_styling()
```

:::

::::

. . . 

```{r}
#| message: false
#| warning: false
#| fig-align: center

point_colors <- c("confidence" = "purple")

ggplot() +
  geom_line(data = new_data,
            aes(x = flipper_length_mm, 
                y = flip_pred),
            color = "blue",
            lwd = 2,
            alpha = 0.5) + 
  geom_point(data = penguins_train,
             aes(x = flipper_length_mm,
                 y = body_mass_g),
             color = "black",
             alpha = 0.4) +
  geom_errorbar(data = new_penguin,
                aes(x = flipper_length_mm,
                    ymin = .conf_lower,
                    ymax = .conf_upper,
                    color = "confidence"),
                alpha = 0.75,
                lwd = 1.5,
                width = 6) +
  geom_point(data = new_penguin,
             aes(x = flipper_length_mm,
                 y = .pred),
             color = "darkgreen",
             size = 6) + 
  scale_color_manual(values = point_colors) + 
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)",
       color = "Interval") 
```


## Using the Model to Make Predictions

What is <font color="green">the body mass of a penguin whose flipper length is 212mm</font>?

+ Somewhere between `r round(new_penguin %>% pull(.pred_lower), 1)`g and `r round(new_penguin %>% pull(.pred_upper), 1)`g, with 95% confidence.

What is <font color="purple">the *average* body mass of *all* penguins whose flipper length is 212mm</font>?

+ Somewhere between `r round(new_penguin %>% pull(.conf_lower), 1)`g and `r round(new_penguin %>% pull(.conf_upper), 1)`g, with 95% confidence.

. . . 

```{r}
#| message: false
#| warning: false
#| fig-align: center

point_colors <- c("prediction" = "darkgreen", "confidence" = "purple")

ggplot() +
  geom_line(data = new_data,
            aes(x = flipper_length_mm, 
                y = flip_pred),
            color = "blue",
            lwd = 2,
            alpha = 0.5) + 
  geom_point(data = penguins_train,
             aes(x = flipper_length_mm,
                 y = body_mass_g),
             color = "black",
             alpha = 0.4) +
  geom_errorbar(data = new_penguin,
                aes(x = flipper_length_mm,
                    ymin = .pred_lower,
                    ymax = .pred_upper,
                    color = "prediction"),
                alpha = 0.5,
                lwd = 1.5,
                width = 5) +
  geom_errorbar(data = new_penguin,
                aes(x = flipper_length_mm,
                    ymin = .conf_lower,
                    ymax = .conf_upper,
                    color = "confidence"),
                alpha = 1,
                lwd = 1.5,
                width = 5) +
  geom_point(data = new_penguin,
             aes(x = flipper_length_mm,
                 y = .pred),
             color = "darkgreen",
             size = 6) + 
  scale_color_manual(values = point_colors) + 
  labs(x = "Flipper Length (mm)",
       y = "Body Mass (g)",
       color = "Interval")
```

## Using the Model to Make Predictions

. . .

<center> 

Let's try this!

</center>

1. Write down two predictive questions associated with Air BnB rental prices that could be answered with your model
2. Can you differentiate versions of those questions that could be answered by *confidence* versus *prediction* intervals? What is the difference?
3. Compute your model's prediction for a particular value of the predictor you chose
4. Similarly, compute the lower and upper bounds for bound confidence and prediction intervals for your chosen value
5. **(Challenge)** Can you plot the result?
6. Interpret your model's predictions

    + Clearly differentiate the model prediction, versus the bounds for the confidence interval, versus the bounds for the prediction interval

## Summary

+ Simple linear regression models are models of the form $\displaystyle{\mathbb{E}\left[y\right] = \beta_0 + \beta_1 x}$, with one predictor.
+ We assume that the sole predictor is linearly associated with the response, independent of any other features. 
+ The *global test for model utility* and the *test for significance of the predictor* are identical in this case.
+ We further assess simple linear regression models with summary metrics like $R^2_{\text{adj}}$, RMSE (both training and testing), as well as residual plots.
+ The intercept is the expected response when the predictor takes a value of 0, which may not be meaningful or supported.
+ The coefficient on the predictor can be interpreted as a slope.
+ We can predict responses for a single observation, or an average over all observations having the same value of the predictor.



