---
title: 'Homework Problems: Introduction to R'
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: lab.css
    keep-md: true
---

```{r global-options, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)

```

This homework set is intended to provide you with practice manipulating and visualizing data. 

**Feel free to ask questions or for help with errors in the Slack channel or during our weekly meetings.** 

For these problems, we will use the `nycflights13` datasets, which can be downloaded using the commands

```{r echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE}
install.packages("nycflights13")
library(nycflights13)
```

```{r echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
library(nycflights13)
```

You now have access to five data frames: `airlines`, `airports`, `flights`, `weather`, and `planes`. A portion of the `flights` dataframe is shown below:

```{r echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE}
flights
```

You can read more about what each of the variables represents by running `?flights` in the console.

1. Visualize the departure delays in the `flights` data set using a histogram. You should relabel the axes and give your plot a title. What information have you learned from the plot?

2. Create a new column in the data set called `time_gained` which is equal to the departure delay subtracted from the arrival delay. Visualize the `time_gained` variable using a box plot. Describe what you see.

3. Do some airlines tend to have more arrival delays than others? Use side-by-side box plots to visualize the arrival delays of each airline.

4. Let's answer the question above in a slightly different way. Calculate the median arrival delays of each airline using the `median()` function. (You'll want to add the argument `na.rm = TRUE` inside the `median()` function to ignore the `NA` values in the data frame.) Sort the airline carriers from the airlines by median arrival delay. Which airlines are consistently ahead of schedule? You can consult the `airlines` data frame to find the airline names that correspond to the abbreviations. Even better, you can use `left_join()` to join the `airlines` data to your data frame of median arrival delays!

5. From the `weather` data frame, visualize the temperature at the `EWR` airport during the entire month of January in 2013. *Hint:* Use the `origin`, `year`, and `month` variables to `filter()` the data frame, and use `time_hour` as the x coordinate in the plot.

6. Make a bar plot which visualizes the top 30 destinations for planes which depart from `JFK` airport.

7. Pick some other variables that you are interested in from these data frames and make appropriate visualizations.
