---
title: "Student Notes Template (Replace Title With Day or Section Number)"
author: "Student Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
#We'll load the tidyverse since it includes the majority of packages we'll want
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(kableExtra)
```

**General Note:** You should keep a copy of this template on your computer. Open it at the beginning of each class period, and use `File -> Save As` to save a new copy corresponding to that class period. This way you'll have a notebook for each individual class meeting and you won't ever overwrite the template file or your previous notes, on accident.

Add your own description of the topic...You can type text freely outside of the code chunks.

## Read in Data

We'll start the semester with pre-built notes utilizing the palmer penguins data. Since those notes are pre-built with all of the necessary code and output, I think its a good idea to rewrite your own notes using a different data set. I'm reading in a *TidyTuesday* dataset on rental properties in the San Francisco area. You are certainly free to choose a different dataset if you'd like though! You can even switch it up from class meeting to class meeting if you are feeling adventurous.

```{r}
#Penguins Data
penguins <- penguins

#SF Rentals Data
rent <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-05/rent.csv')
#permits <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-05/sf_permits.csv')
#new_construction <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-07-05/new_construction.csv')

rent %>% 
  head() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("hover", "striped"))
```

## 
